---
title: "Moonmilk_SeqCode_Manuscript_080923_ARS"
author: "André Rodrigues Soares"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggrepel)
library(kableExtra)
```

```{r}
bac_phyla_of_interest = ""
bac_classes_of_interest = c("Terriglobia", "UBA10030", "RBG-16-71-46", "UBA2214", 
                            "Krumholzibacteria", "Methylomirabilia", "Nanoarchaeia", 
                            "ABY1", "Nitrososphaeria", "MSB-5A5")
```

Only MAGs >= 70% completeness & <= 10% contamination were kept.

```{r message=FALSE, warning=FALSE}
hq_mags_mg1 = read_tsv("mgshot_S7938Nr1_DAS_Tool_uBin_curated_checkm2_gt70lt10_IDs", 
                   col_names = "ID")$ID
hq_mags_mg2 = read_tsv("mgshot_S7938Nr2_DAS_Tool_uBin_curated_checkm2_gt70lt10_IDs", 
                   col_names = "ID")$ID
```

### Summary stats and taxonomy of MAGs of interest:

```{r message=FALSE, warning=FALSE}
checkm2_mg1 = read_tsv("mgshot_S7938Nr1_quality_report.tsv") %>% 
  filter(Name %in% hq_mags_mg1) %>% 
  select(Name, Completeness, Contamination, Genome_Size, GC_Content) %>% 
  add_column(sample = "mgshot_S7938Nr1") %>% 
  filter(Completeness >= 70 & Contamination <= 10)

checkm2_mg2 = read_tsv("mgshot_S7938Nr2_quality_report.tsv") %>% 
  filter(Name %in% hq_mags_mg2) %>% 
  select(Name, Completeness, Contamination, Genome_Size, GC_Content) %>% 
  add_column(sample = "mgshot_S7938Nr2")

checkm2_all = bind_rows(checkm2_mg1, checkm2_mg2) %>% 
  filter(Completeness >= 70 & Contamination <= 10)
```

```{r message=FALSE, warning=FALSE}
gtdbtk_mg1_bac = read_tsv("mgshot_S7938Nr1_gtdbtk.bac120.summary.tsv") %>% 
  filter(user_genome %in% hq_mags_mg1) %>% 
  select(user_genome, classification) %>% 
  add_column(sample = "mgshot_S7938Nr1")
gtdbtk_mg1_arc = read_tsv("mgshot_S7938Nr1_gtdbtk.ar53.summary.tsv") %>% 
  filter(user_genome %in% hq_mags_mg1) %>% 
  select(user_genome, classification) %>% 
  add_column(sample = "mgshot_S7938Nr1")

gtdbtk_mg2_bac = read_tsv("mgshot_S7938Nr2_gtdbtk.bac120.summary.tsv") %>% 
  filter(user_genome %in% hq_mags_mg2) %>% 
  select(user_genome, classification) %>% 
  add_column(sample = "mgshot_S7938Nr2")
gtdbtk_mg2_arc = read_tsv("mgshot_S7938Nr2_gtdbtk.ar53.summary.tsv") %>% 
  filter(user_genome %in% hq_mags_mg2) %>% 
  select(user_genome, classification) %>% 
  add_column(sample = "mgshot_S7938Nr2")

gtdbtk_all = bind_rows(gtdbtk_mg1_bac, gtdbtk_mg1_arc,
                       gtdbtk_mg2_bac, gtdbtk_mg2_arc) %>% 
  rename(Name = user_genome) %>% 
  mutate(classification = str_remove_all(classification,
                                     paste(c("d__","p__","c__",
                                             "o__","f__","g__"), 
                                           collapse = "|"))) %>% 
  separate(classification, into = c("domain", "phylum", "class",
                                    "order", "family", "genus"), sep = ";") %>% 
  filter(phylum %in% bac_phyla_of_interest | class %in% bac_classes_of_interest)
```

```{r}
abund_list = list.files(".", pattern = "_abunds.txt")
mag_abunds = read_tsv(abund_list, id = "sample", 
                      col_names = c("Name","rel_abund","mean_cov"), skip = 1) %>% 
  # filter(Name %in% mags_of_interest) %>% 
  mutate(sample = str_remove(sample, "_coverm_mag_abunds.txt"),
         rel_abund = rel_abund)

all_checkm_gtdbtk_abunds = left_join(gtdbtk_all, checkm2_all,
                              by = c("Name", "sample")) %>% 
                    left_join(mag_abunds,
                              by = c("Name", "sample"))

mags_of_interest = all_checkm_gtdbtk_abunds$Name

```

```{r}
all_checkm_gtdbtk_abunds %>% 
  select(-domain) %>% 
  arrange(phylum, desc(Completeness), Contamination) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_classic(full_width = F)
```

### Metabolisms as per DRAM

```{r}
po_order = rev(c("Thermoproteota, Nitrososphaeria, Nitrososphaerales", "Methylomirabilota, Methylomirabilia, Methylomirabilales",
             "Zixibacteria, MSB-5A5, GN15", "Eisenbacteria, RBG-16-71-46, RBG-16-71-46", "KSB1, UBA2214, DRLW01",
             "Acidobacteriota, Terriglobia, Bryobacterales", "Acidobacteriota, Terriglobia, JAJPHL01",
             "Krumholzibacteriota, Krumholzibacteria, G020349025", "Krumholzibacteriota, Krumholzibacteria, WVZY01",
             "Bacteroidota, UBA10030, UBA10030"))

dram_mg1_sum = read_tsv("mgshot_S7938Nr1_DAS_Tool_uBin_curated_DRAM_out_distill/product.tsv") %>% 
  gather("metabolism", "values", -genome) %>% 
  filter(genome %in% mags_of_interest & values > 0.6) %>% 
  add_column(sample = "mgshot_S7938Nr1")

dram_mg2_sum = read_tsv("mgshot_S7938Nr2_DAS_Tool_uBin_curated_DRAM_out_distill/product.tsv") %>% 
  gather("metabolism", "values", -genome) %>% 
  filter(genome %in% mags_of_interest & values > 0.6) %>% 
  add_column(sample = "mgshot_S7938Nr2")

dram_mg1_mg2_sum = bind_rows(dram_mg1_sum, dram_mg2_sum) %>% 
  rename(Name = genome)

dram_mg1_mg2_sum_all_data = left_join(dram_mg1_mg2_sum,
                                     all_checkm_gtdbtk_abunds,
                                     by = c("Name", "sample"))

# ggplot(dram_mg1_mg2_sum_all_data,
#        aes(Name, metabolism, fill = values)) +
#   geom_tile(colour = 'black') +
#   facet_grid(.~phylum+class, 
#              scales = "free_x", space = "free_x") +
#   scale_fill_viridis_c() +
#   theme_bw() +
#   theme(axis.text.x = element_blank(),
#         axis.ticks.x = element_blank(),
#         axis.title = element_blank())

order_met = rev(c("Citrate cycle (TCA cycle, Krebs cycle)", 
              "Entner-Doudoroff pathway, glucose-6P => glyceraldehyde-3P + pyruvate", 
              "Glycolysis (Embden-Meyerhof pathway), glucose => pyruvate", "Glyoxylate cycle",
              "Pentose phosphate pathway (Pentose phosphate cycle)", "Reductive citrate cycle (Arnon-Buchanan cycle)",
              "Reductive pentose phosphate cycle (Calvin cycle)",
              "Complex I: NADH:quinone oxidoreductase, prokaryotes", "Complex II: Fumarate reductase, prokaryotes", 
              "Complex II: Succinate dehydrogenase, prokaryotes", "Complex III: Cytochrome bd ubiquinol oxidase", 
              "Complex IV High affinity: Cytochrome bd ubiquinol oxidase", 
              "Complex IV Low affinity: Cytochrome c oxidase, prokaryotes", 
              "Complex V: F-type ATPase, prokaryotes and chloroplasts", "Complex V: V/A-type ATPase, prokaryotes", 
              "Methanogenesis and methanotrophy: acetate => methane, pt 1", 
              "Methanogenesis and methanotrophy: acetate => methane, pt 2", 
              "Methanogenesis and methanotrophy: acetate => methane, pt 3", 
              "Methanogenesis and methanotrophy: methane => methanol, with oxygen (pmo)",
              "Methanogenesis and methanotrophy: putative but not defining CO2 => methane", 
              "Nitrogen metabolism: Bacterial/Archaeal ammonia oxidation", 
              "Nitrogen metabolism: Dissimilatory nitrite reduction to ammonia (DNRA)", 
              "Nitrogen metabolism: nitrite => nitric oxide", 
              "Nitrogen metabolism: nitrous oxide => nitrogen", 
              "SCFA and alcohol conversions: acetate, pt 1", "SCFA and alcohol conversions: Alcohol production", 
              "SCFA and alcohol conversions: lactate D", "SCFA and alcohol conversions: pyruvate => acetyl CoA v1", 
              "SCFA and alcohol conversions: pyruvate => acetyl CoA v2",
              "CAZy: Amorphous Cellulose", "CAZy: Arabinan", "CAZy: Mixed-Linkage glucans", 
              "CAZy: Polyphenolics", "CAZy: Xyloglucan", 
              "Other Reductases: arsenate reduction, pt 1"))


p_amm = ggplot(dram_mg1_mg2_sum_all_data %>% 
            mutate(po = paste0(phylum, ", ",class, ", ", order)),
       aes(fct_relevel(metabolism, order_met), Name, fill = values*100)) +
  geom_tile(colour = 'black') +
  geom_vline(xintercept = c(7.5, 15.5, 20.5, 24.5, 29.5)) +
  facet_grid(fct_relevel(po, rev(po_order))~., 
             scales = "free", space = "free") +
  scale_fill_viridis_c() +
  scale_x_discrete(position = "top", 
                   limits = rev, expand = c(0,0)) +
  guides(fill = guide_colorbar(title = "Pathway Completeness (%)")) +
  theme_bw() +
  theme(axis.text.x.top = element_text(angle = 30, hjust = 0, vjust = 0),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title = element_blank(),
        strip.text.y = element_text(angle = 0, face = "bold", size = 8),
        strip.background = element_rect(fill = "white"),
        legend.position = "bottom")

ggsave("AMM_DRAM_metabolisms.png", height = 8, width = 13)

cc = dram_mg1_mg2_sum_all_data %>% 
            mutate(po = paste0(phylum, ", ",class, ", ", order)) %>% 
            select(po, Completeness, Contamination, Name) %>% 
            gather("keys", "values", -po, -Name) %>% 
            ggplot(aes(keys, Name, 
                           fill = values,
                           label = round(values, 1))) +
  geom_tile(colour = "black") +
  geom_text(colour = "white", size = 2) +
  facet_grid(fct_relevel(po, rev(po_order))~., 
             scales = "free", space = "free") +
  scale_fill_viridis_c(trans = "sqrt", option = "turbo",
                       direction = -1,
                       breaks = c(5, 10, 25, 50, 70, 100),
                       labels = c(5, 10, 25, 50, 70, 100)) +
  scale_x_discrete(expand = c(0,0)) +
  guides(fill = guide_colorbar(title = "MAG Completeness & Contamination (%)")) +
  theme_bw() +
  theme(axis.text.x.bottom = element_text(angle = 90, 
                                          hjust = 1, vjust = 1),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title = element_blank(),
        strip.text.y = element_blank(),
        strip.background = element_rect(fill = "white"),
        legend.position = "bottom")

cc + p_amm +
  plot_layout(widths = c(0.05, 1), guides = "collect") & theme(legend.position = 'bottom')

ggsave("AMM_DRAM_metabolisms_wcc.png", height = 9, width = 13)
```

```{r}
ggplot(dram_mg1_mg2_sum_all_data %>% 
            mutate(po = paste0(phylum, ",\n",class, ",\n", order)),
       aes(Name, metabolism, fill = values)) +
  geom_tile(colour = 'black') +
  facet_grid(.~fct_relevel(po, rev(po_order)), 
             scales = "free") +
  scale_fill_viridis_c() +
  scale_x_discrete(position = "top", 
                   limits = rev, expand = c(0,0)) +
  theme_bw() +
  theme(axis.text.x.top = element_blank(),
        axis.ticks.x.top = element_blank(),
        axis.title = element_blank(),
        strip.text.x = element_text(angle = 0, 
                                    face = "bold", size = 7),
        strip.background = element_rect(fill = "white"))

# ggsave("AMM_DRAM_metabolisms.png", height = 8, width = 13)
```

### defense-finder

```{r}
mg1_df_stats_files = list.files("defensefinder/mgshot_S7938Nr1", 
                          recursive = T, full.names = T,
                          pattern = "defense_finder_hmmer.tsv")

mg1_df_stats = read_tsv(mg1_df_stats_files, id = "file_name") %>% 
  filter(i_eval >= 1e-5)

mg1_df_files = list.files("defensefinder/mgshot_S7938Nr1", 
                          recursive = T, full.names = T,
                          pattern = "defense_finder_systems.tsv")

mg1_df = read_tsv(mg1_df_files, id = "file_name") %>% 
  mutate(clean_name = str_remove(file_name, "defensefinder/mgshot_S7938Nr1/")) %>% 
  mutate(clean_name = str_remove(clean_name, "_out/defense_finder_systems.tsv")) %>% 
  add_column(sample = "mgshot_S7938Nr1") %>% 
  filter(!protein_in_syst %in% mg1_df_stats$hit_id) %>% 
  select(type, subtype, genes_count, clean_name, sample)

mg2_df_stats_files = list.files("defensefinder/mgshot_S7938Nr2", 
                          recursive = T, full.names = T,
                          pattern = "defense_finder_hmmer.tsv")

mg2_df_stats = read_tsv(mg2_df_stats_files, id = "file_name") %>% 
  filter(i_eval >= 1e-5)

mg2_df_files = list.files("defensefinder/mgshot_S7938Nr2", 
                          recursive = T, full.names = T,
                          pattern = "defense_finder_systems.tsv")

mg2_df = read_tsv(mg2_df_files, id = "file_name") %>% 
  mutate(clean_name = str_remove(file_name, "defensefinder/mgshot_S7938Nr2/")) %>% 
  mutate(clean_name = str_remove(clean_name, "_out/defense_finder_systems.tsv")) %>% 
  add_column(sample = "mgshot_S7938Nr2") %>% 
  filter(!protein_in_syst %in% mg2_df_stats$hit_id) %>% 
  select(type, subtype, genes_count, clean_name, sample)

mg3_df_stats_files = list.files("defensefinder/mgshot_S7938Nr3", 
                          recursive = T, full.names = T,
                          pattern = "defense_finder_hmmer.tsv")

mg3_df_stats = read_tsv(mg3_df_stats_files, id = "file_name") %>% 
  filter(i_eval >= 1e-5)

mg3_df_files = list.files("defensefinder/mgshot_S7938Nr3", 
                          recursive = T, full.names = T,
                          pattern = "defense_finder_systems.tsv")

mg3_df = read_tsv(mg3_df_files, id = "file_name") %>% 
  mutate(clean_name = str_remove(file_name, "defensefinder/mgshot_S7938Nr3/")) %>% 
  mutate(clean_name = str_remove(clean_name, "_out/defense_finder_systems.tsv")) %>% 
  add_column(sample = "mgshot_S7938Nr3") %>% 
  filter(!protein_in_syst %in% mg3_df_stats$hit_id) %>% 
  select(type, subtype, genes_count, clean_name, sample)

all_df = bind_rows(mg1_df, mg2_df, mg3_df) %>% 
  rename(Name = clean_name) %>% 
  left_join(left_join(all_checkm_gtdbtk_abunds, mag_abunds, by = c("Name", "sample")), by = c("Name", "sample")) %>% 
  filter(Name %in% mags_of_interest)
```

Diversity of phage and plasmid defense systems for active moonmilk taxa

```{r}
df = ggplot(all_df %>% 
         mutate(po = paste0(phylum, ", ",class, ", ", order)) %>% 
         group_by(subtype, po) %>% 
         summarise(count = sum(genes_count),
                   Name = Name) %>% 
         filter(count > 1 & !grepl("CAS", subtype)) %>% 
         distinct(),
       aes(subtype, Name, fill = count, label = count)) +
  geom_tile(colour = 'black') +
  geom_text(aes(colour = ifelse(count > 12, "black", "white"))) +
  facet_grid(fct_relevel(po, rev(po_order)) ~ .,
             scales = "free", space = "free") +
  scale_fill_viridis_c() +
  scale_colour_identity() +
  scale_x_discrete(expand = c(0,0)) +
  scale_y_discrete(expand = c(0,0)) +
  guides(fill = guide_colorbar(title = "Number of copies found per MAG")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title = element_blank(),
        strip.text.y = element_text(angle = 0, face = "bold"),
        strip.background = element_rect(fill = "white"),
        legend.position = "bottom")
df
ggsave("AMM_defense_sytems.png", height = 6, width = 11)

crispr_mags = readxl::read_xlsx("../virus_host_matching.a/Genomes_repeat_crisprsubtype_probablility.xlsx",
                  col_names = c("Name", "repeats", "subtype", "prob")) %>% 
  filter(!grepl("../", Name)) %>% 
  mutate(Name = str_remove(Name, ".fasta")) %>% 
  full_join(all_checkm_gtdbtk_abunds, by = "Name") %>% 
  filter(Name %in% mags_of_interest)

crispr_spls = readxl::read_xlsx("../virus_host_matching.a/Genomes_repeat_crisprsubtype_probablility.xlsx",
                  col_names = c("Name", "repeats", "subtype", "prob")) %>% 
  filter(!grepl("../", Name)) %>% 
  mutate(Name = str_remove(Name, ".fasta"),
         sample = str_extract(Name, "^[^_]+_[^_]+")) %>% 
  filter(prob >= 0.75) %>% 
  group_by(sample, subtype) %>% 
  summarise(count = n())

crispr_spls %>% 
  pivot_wider(names_from = "subtype", values_from = "count") %>% 
  replace(is.na(.), 0) %>% 
  mutate(sample = str_replace(sample, "mgshot_S7938Nr1", "mgshot_S7938Nr1 (AMM_F)"),
         sample = str_replace(sample, "mgshot_S7938Nr2", "mgshot_S7938Nr2 (AMM_B)"),
         sample = str_replace(sample, "mgshot_S7938Nr3", "mgshot_S7938Nr3 (DMM_F)")) %>% 
  kable() %>% 
  kable_paper() %>%
  kable_styling(full_width = F,
                bootstrap_options = c("striped", "condensed"))
```

```{r}
cm = ggplot(crispr_mags %>% 
         mutate(po = paste0(phylum, ", ",class, ", ", order)) %>% 
         group_by(subtype, po) %>% 
         summarise(count = n(),
                   Name = Name,
                   prob = prob) %>% 
          filter(Name %in% df$data$Name),
       aes(subtype, Name, fill = count, label = count)) +
  geom_tile(aes(alpha = prob),
            colour = 'black') +
  geom_text(aes(colour = ifelse(prob < .75, "black", "white"))) +
  facet_grid(fct_relevel(po, rev(po_order)) ~ .,
             scales = "free", space = "free") +
  scale_fill_viridis_c() +
  scale_colour_identity() +
  scale_x_discrete(expand = c(0,0), 
                   limits = c("I-A", "I-C", "II-B", "III-B", "IV-A2")) +
  scale_y_discrete(expand = c(0,0)) +
  guides(fill = guide_colorbar(title = "Number of copies found per MAG")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
        # axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title = element_blank(),
        strip.text.y = element_blank(),
        # strip.text.y = element_text(angle = 0, face = "bold"),
        strip.background = element_rect(fill = "white"),
        legend.position = "none")

cm + df +
  plot_layout(widths = c(0.2, 1))

ggsave("AMM_defense_sytems_ccf.png", height = 6, width = 14)
```

```{r}
ggplot(all_df %>% 
         mutate(po = paste0(phylum, ", ", order)) %>% 
         group_by(subtype, po, sample) %>% 
         summarise(count = sum(genes_count),
                   Name = Name,
                   phylum = phylum,
                   order = order) %>% 
         group_by(Name, sample) %>% 
         mutate(n = n(),
                norm_count = round(count/n, 1)) %>% 
         group_by(po, subtype, sample) %>% 
         summarise(mean_norm_count = round(mean(norm_count), 1),
                   subtype = subtype) %>% 
         filter(mean_norm_count >= 1), 
       aes(po, subtype, fill = mean_norm_count, label = mean_norm_count)) +
  geom_tile(colour = 'black') +
  geom_text(aes(colour = ifelse(mean_norm_count > 30, "black", "white"))) +
  facet_grid(sample~., 
             scales = "free", space = "free") +
  scale_fill_viridis_c() +
  scale_colour_identity() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
        legend.position = "none")
```

## METABOLIC hits for AMM MAGs of interest

```{r}
metabolic_mg1 = read_tsv("mgshot_S7938Nr1_DAS_Tool_uBin_curated_METABOLIC/MW-score_result/MW-score_result.txt") %>% 
  gather("Name", "MW", -Function, -`MW-score for each function`) %>% 
  add_column("sample" = "mgshot_S7938Nr1")

metabolic_mg2 = read_tsv("mgshot_S7938Nr2_DAS_Tool_uBin_curated_METABOLIC/MW-score_result/MW-score_result.txt") %>% 
  gather("Name", "MW", -Function, -`MW-score for each function`) %>% 
  add_column("sample" = "mgshot_S7938Nr1")

# metabolic_mg3 = read_tsv("mgshot_S7938Nr3_DAS_Tool_uBin_curated_METABOLIC/MW-score_result/MW-score_result.txt") %>% 
#   gather("Name", "MW", -Function, -`MW-score for each function`) %>% 
#   add_column("sample" = "mgshot_S7938Nr1")

all_metabolic = bind_rows(metabolic_mg1, metabolic_mg2) %>% 
  right_join(all_checkm_gtdbtk_abunds, 
            by = c("Name", "sample")) %>% 
  filter(phylum %in% bac_phyla_of_interest | class %in% bac_classes_of_interest) %>% 
  filter(!is.na(Function))
```

```{r}
ggplot(all_metabolic %>% 
            mutate(po = paste0(phylum, ", ",class, ", ", order)),
       aes(Function, Name, fill = MW)) +
  geom_tile(colour = 'black') +
  facet_grid(fct_relevel(po, rev(po_order))~., 
             scales = "free", space = "free") +
  scale_fill_viridis_c(trans = "sqrt") +
  scale_x_discrete(position = "top", 
                   limits = rev, expand = c(0,0)) +
  theme_bw() +
  theme(axis.text.x.top = element_text(angle = 30, hjust = 0, vjust = 0),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title = element_blank(),
        strip.text.y = element_text(angle = 0, face = "bold"),
        strip.background = element_rect(fill = "white"))

```

```{r}
metabolic_mg1_comp = readxl::read_xlsx("mgshot_S7938Nr1_DAS_Tool_uBin_curated_METABOLIC/METABOLIC_result.xlsx",
                                       sheet = 4) %>% 
  group_by(Module) %>% 
  mutate(Module.step.total = n()) %>% 
  gather("Name", "presence_absence", -Module.step, -Module, -KO.id, -Module.Category, -Module.step.total) %>% 
  mutate(presence_absence_bin = case_when(presence_absence == "Present" ~ "1",
                                      TRUE ~ "0")) %>% 
  mutate(presence_absence_bin = as.numeric(as.character(presence_absence_bin))) %>% 
  group_by(Module, Name) %>% 
  mutate(step_sum = sum(presence_absence_bin),
         completeness = step_sum/Module.step.total) %>% 
  select(Module, Module.Category, Name, completeness) %>% 
  distinct() %>% 
  add_column("sample" = "mgshot_S7938Nr1")

metabolic_mg2_comp = readxl::read_xlsx("mgshot_S7938Nr2_DAS_Tool_uBin_curated_METABOLIC/METABOLIC_result.xlsx",
                                       sheet = 4) %>% 
  group_by(Module) %>% 
  mutate(Module.step.total = n()) %>% 
  gather("Name", "presence_absence", -Module.step, -Module, -KO.id, -Module.Category, -Module.step.total) %>% 
  mutate(presence_absence_bin = case_when(presence_absence == "Present" ~ "1",
                                      TRUE ~ "0")) %>% 
  mutate(presence_absence_bin = as.numeric(as.character(presence_absence_bin))) %>% 
  group_by(Module, Name) %>% 
  mutate(step_sum = sum(presence_absence_bin),
         completeness = step_sum/Module.step.total) %>% 
  select(Module, Module.Category, Name, completeness) %>% 
  distinct() %>% 
  add_column("sample" = "mgshot_S7938Nr2")

metabolic_all_comp = bind_rows(metabolic_mg1_comp, metabolic_mg2_comp) %>% 
  mutate(Name = str_remove(Name, ".Module.step.presence")) %>% 
  right_join(all_checkm_gtdbtk_abunds, 
            by = c("Name", "sample")) %>% 
  filter(phylum %in% bac_phyla_of_interest | class %in% bac_classes_of_interest)

```

```{r}
ggplot(metabolic_all_comp %>% 
            mutate(po = paste0(phylum, ", ",class, ", ", order)) %>% 
            filter(completeness >= 0.7),
       aes(Module, Name, fill = completeness)) +
  geom_tile(colour = 'black') +
  facet_grid(fct_relevel(po, rev(po_order))~Module.Category, 
             scales = "free", space = "free") +
  scale_fill_viridis_c(trans = "log10") +
  scale_x_discrete(position = "top", 
                   limits = rev, expand = c(0,0)) +
  theme_bw() +
  theme(axis.text.x.top = element_text(angle = 30, hjust = 0, vjust = 0),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title = element_blank(),
        strip.text.y = element_text(angle = 0, face = "bold"),
        strip.background = element_rect(fill = "white"))
```

## KSB1, MSB-5A5 DRAM hits

```{r}
ksb1_checkm2 = read_tsv("KSB1_checkm2_291123_quality_report.tsv") %>% 
  select(Name, Completeness, Contamination, Genome_Size, GC_Content) %>% 
  filter(Completeness >= 70 & Contamination <= 10)

msb_checkm2 = read_tsv("MSB-5A5_checkm2_291123_quality_report.tsv") %>% 
  select(Name, Completeness, Contamination, Genome_Size, GC_Content) %>% 
  filter(Completeness >= 70 & Contamination <= 10)

ksb1_gtdb = read_tsv("KSB1_gtdbtk.bac120.summary.tsv") %>% 
  rename(Name = user_genome) %>% 
  mutate(classification = str_remove_all(classification,
                                     paste(c("d__","p__","c__",
                                             "o__","f__","g__"), 
                                           collapse = "|"))) %>% 
  separate(classification, into = c("domain", "phylum", "class",
                                    "order", "family", "genus"), sep = ";") %>% 
  select(Name, domain, phylum, class, order, family, genus)

msb_gtdb = read_tsv("MSB-5A5_gtdbtk.bac120.summary.tsv") %>% 
  rename(Name = user_genome) %>% 
  mutate(classification = str_remove_all(classification,
                                     paste(c("d__","p__","c__",
                                             "o__","f__","g__"), 
                                           collapse = "|"))) %>% 
  separate(classification, into = c("domain", "phylum", "class",
                                    "order", "family", "genus"), sep = ";") %>% 
  select(Name, domain, phylum, class, order, family, genus)

ksb1_energy = readxl::read_xlsx("KSB1_DRAM_291123_distill//metabolism_summary.xlsx", 
                             sheet = 4) %>% 
  group_by(module) %>% 
  mutate(count = n()) %>% 
  gather("Name", "values", -gene_id, -gene_description, -module, -header, -subheader, -count) %>% 
  mutate(values = as.double(as.character(values))) %>%
  left_join(ksb1_checkm2, by = "Name")

msb_energy = readxl::read_xlsx("MSB-5A5_DRAM_291123_distill//metabolism_summary.xlsx", 
                             sheet = 4) %>% 
  group_by(module) %>% 
  mutate(count = n()) %>% 
  gather("Name", "values", -gene_id, -gene_description, -module, -header, -subheader, -count) %>% 
  mutate(values = as.double(as.character(values))) %>%
  left_join(msb_checkm2, by = "Name")

ksb1_energy_summary = ksb1_energy %>% 
  add_column(type = "energy") %>% 
  group_by(module, Name) %>% 
  mutate(hits = case_when(values >= 1 ~ 1,
                          TRUE ~ 0),
    completeness = sum(hits)/count) %>% 
  mutate(completeness_na = case_when(hits >= 1 ~ completeness,
                                     TRUE ~ 0))

msb_energy_summary = msb_energy %>% 
  add_column(type = "energy") %>% 
  group_by(module, Name) %>% 
  mutate(hits = case_when(values >= 1 ~ 1,
                          TRUE ~ 0),
    completeness = sum(hits)/count) %>% 
  mutate(completeness_na = case_when(hits >= 1 ~ completeness,
                                     TRUE ~ 0))

ksb1_energy_summary_plot = ksb1_energy_summary %>% 
  filter(!grepl("Ribosome", module, ignore.case= T) & 
         !grepl("polymerase", module, ignore.case = T)) %>% 
  group_by(Name, gene_description) %>% 
  reframe(hits = sum(values),
          header = header,
          module = module,
          # phylum = phylum,
          # class = class,
          # order = order,
          # genus = genus,
          completeness = completeness,
          completeness_na = completeness_na) %>% 
  left_join(ksb1_gtdb, by = "Name")

msb_energy_summary_plot = msb_energy_summary %>% 
  filter(!grepl("Ribosome", module, ignore.case= T) & 
         !grepl("polymerase", module, ignore.case = T)) %>% 
  group_by(Name, gene_description) %>% 
  reframe(hits = sum(values),
          header = header,
          module = module,
          # phylum = phylum,
          # class = class,
          # order = order,
          # genus = genus,
          completeness = completeness,
          completeness_na = completeness_na) %>% 
  left_join(msb_gtdb, by = "Name")

carbon_pathways = c("Reductive citrate cycle (Arnon-Buchanan cycle)", "Dicarboxylate-hydroxybutyrate cycle", "Hydroxypropionate-hydroxybutylate cycle", 
                    "Methanogenesis, acetate => methane", "Malonate semialdehyde pathway, propanoyl-CoA => acetyl-CoA", "Acetyl-CoA pathway", 
                    "Methanogenesis, CO2 => methane",  "Reductive pentose phosphate cycle (Calvin cycle)", 
                    "Reductive pentose phosphate cycle, glyceraldehyde-3P => ribulose-5P", 
                    "Reductive pentose phosphate cycle, ribulose-5P => glyceraldehyde-3P",
                    "Reductive acetyl-CoA pathway (Wood-Ljungdahl pathway)", "Dissimilatory nitrate reduction, nitrate => ammonia", 
                    "Succinate dehydrogenase, prokaryotes", "Methanogenesis, methanol => methane", 
                    "Methanogenesis, methylamine/dimethylamine/trimethylamine => methane", "Assimilatory sulfate reduction, sulfate => H2S",
                    "Dissimilatory sulfate reduction, sulfate => H2S", "3-Hydroxypropionate bi-cycle", 
                    "Complete nitrification, comammox, ammonia => nitrite => nitrate", "Nitrification, ammonia => nitrite",
                    "Arsenate", "Methane oxidation, methanotroph, methane => formaldehyde", "nitrite + ammonia => nitrogen", 
                    "Denitrification, nitrate => nitrogen", "hydrogenase", "Assimilatory nitrate reduction, nitrate => ammonia")


ksb1_energy_summary %>% 
    left_join(ksb1_gtdb, by = "Name") %>% 
    ungroup() %>% 
    filter(!grepl("mgshot", Name)) %>% 
    select(Completeness, Contamination, order, family) %>% 
  ggplot(aes(Contamination, Completeness)) + 
    facet_grid(.~order+family) + 
    geom_hline(yintercept = 80, 
               linetype = "dashed", alpha = .3) +
    geom_vline(xintercept = 5, 
               linetype = "dashed", alpha = .3) +
    geom_point(colour = "gray40",
               size = 3, alpha = 0.5) +
    geom_point(colour = "black",
               shape = 1,
               size = 3, alpha = 0.5) +
    geom_point(data = ksb1_energy_summary %>% 
                        left_join(ksb1_gtdb, by = "Name") %>% 
                        ungroup() %>% 
                        filter(grepl("mgshot", Name)) %>% 
                        select(Completeness, Contamination, order, family),
               mapping = aes(Contamination, Completeness),
               colour = "firebrick3",
               size = 3, alpha = 0.5) +
    theme_bw()

ggsave("KSB1_checkm.png",
       height = 2, width = 20)

ggplot(ksb1_energy_summary_plot %>% 
         filter(module %in% carbon_pathways & 
                  completeness > 0.3 &
                  completeness_na > 0),
       aes(Name, gene_description, fill = completeness_na)) +
  geom_tile(colour = 'black') +
  facet_grid(module~order+family, scales = "free", space = "free_y") +
  scale_fill_viridis_c() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
          strip.background = element_rect(fill = "white"),
          strip.text.y = element_text(face = "bold", angle = 0))

ggsave("KSB1_Cmet.png", 
       height = 14, width = 33)

msb_energy_summary %>% 
    left_join(msb_gtdb, by = "Name") %>% 
    ungroup() %>% 
    filter(!grepl("mgshot", Name)) %>% 
    select(Completeness, Contamination, order, family) %>% 
  ggplot(aes(Contamination, Completeness)) + 
    facet_grid(.~order+family) + 
    geom_hline(yintercept = 80, 
               linetype = "dashed", alpha = .3) +
    geom_vline(xintercept = 5, 
               linetype = "dashed", alpha = .3) +
    geom_point(colour = "gray40",
               size = 3, alpha = 0.5) +
    geom_point(colour = "black",
               shape = 1,
               size = 3, alpha = 0.5) +
    geom_point(data = msb_energy_summary %>% 
                        left_join(msb_gtdb, by = "Name") %>% 
                        ungroup() %>% 
                        filter(grepl("mgshot", Name)) %>% 
                        select(Completeness, Contamination, order, family),
               mapping = aes(Contamination, Completeness),
               colour = "firebrick3",
               size = 3, alpha = 0.5) +
    theme_bw()

ggsave("MSB_checkm.png",
       height = 2, width = 20)

ggplot(msb_energy_summary_plot %>% 
         filter(module %in% carbon_pathways & 
                  completeness > 0.3 &
                  completeness_na > 0),
       aes(Name, gene_description, fill = completeness_na)) +
  geom_tile(colour = 'black') +
  facet_grid(module~order+family, scales = "free", space = "free_y") +
  scale_fill_viridis_c() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
          strip.background = element_rect(fill = "white"),
          strip.text.y = element_text(face = "bold", angle = 0))

ggsave("MSB-5A5_Cmet.png", 
       height = 14, width = 33)
```
